---
title: "DataPreparation"
author: "Sunil Prakash"
date: "2/21/2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preparation and Cleaning Workshop



```{r loadData}

strain <- read.csv("salary-train.csv");
stest <- read.csv("salary-test.csv");


str(strain)

```
### Data Cleaning

```{r}

strain$workclass = as.factor(gsub('?', NA, strain$workclass, fixed = T))
strain$education = as.factor(gsub('?', NA, strain$education, fixed = T))
#strain$material = as.factor(gsub('?', NA, strain$material, fixed = T))
strain$occupation = as.factor(gsub('?', NA, strain$occupation, fixed = T))
strain$relationship = as.factor(gsub('?', NA, strain$relationship, fixed = T))
strain$race = as.factor(gsub('?', NA, strain$race, fixed = T))
strain$sex = as.factor(gsub('?', NA, strain$sex, fixed = T))
strain$native.country = as.factor(gsub('?', NA, strain$native.country, fixed = T))
strain$class = as.factor(gsub('?', NA, strain$class, fixed = T))


stest$workclass = as.factor(gsub('?', NA, stest$workclass, fixed = T))
stest$education = as.factor(gsub('?', NA, stest$education, fixed = T))
#strain$material = as.factor(gsub('?', NA, strain$material, fixed = T))
stest$occupation = as.factor(gsub('?', NA, stest$occupation, fixed = T))
stest$relationship = as.factor(gsub('?', NA, stest$relationship, fixed = T))
stest$race = as.factor(gsub('?', NA, stest$race, fixed = T))
stest$sex = as.factor(gsub('?', NA, stest$sex, fixed = T))
stest$native.country = as.factor(gsub('?', NA, stest$native.country, fixed = T))
stest$class = as.factor(gsub('?', NA, stest$class, fixed = T))

#grep('\\?',strain$workclass)
strain <- na.omit(strain)
stest <- na.omit(stest)

```



### Regression 

```{r}

model <- glm(class ~.,family=binomial(link='logit'),data=strain)

plot(model)


```

```{r}
anova(model, test="Chisq")
```

```{r}
#Testing

fitted.results <- predict(model,newdata=stest[-14],type='response')
fitted.results <- ifelse(fitted.results > 0.5,'>50K.','<=50K.')
xtab <- table(fitted.results,stest$class)

avgAccuracy = sum(diag(xtab)) / sum(xtab)
avgAccuracy



```
the first model provides 84.76% accuracy

Now lets do some more cleaning, looks like two columns capital.gain and capital.loss have many 0 values, which could be result of missing data only,  lets get rid of these columns and try again

```{r}

strain2 <- strain[-10]
strain2 <- strain2[-10]

stest2 <- stest[-10]
stest2 <- stest2[-10]

model2 <- glm(class ~.,family=binomial(link='logit'),data=strain2)

fitted.results2 <- predict(model2,newdata=stest2[-12],type='response')
fitted.results2 <- ifelse(fitted.results2 > 0.5,'>50K.','<=50K.')
xtab <- table(fitted.results2,stest2$class)

avgAccuracy = sum(diag(xtab)) / sum(xtab)
avgAccuracy


```
the accuracy reduces, on removing the two columns to 82.96 %.

lets try doing co-relation analysis and pca to find some  correlated fields  and reduction

```{r}
library(dummies)

new_strain <- dummy.data.frame(strain, names = c("workclass","education",
                                "marital","occupation",
                                "relationship","race","sex","native.country"))
new_stest <- dummy.data.frame(stest, names = c("workclass","education",
                                "marital","occupation",
                                "relationship","race","sex","native.country"))

names(new_strain)


pca <- prcomp(new_strain[-104])
pca_test <- prcomp(new_stest[-103])

screeplot(pca, type = 'l')

```

looks like only 2 pca is enogh for 99% varience

```{r}

strain3 <- strain
stest3 <- stest

strain.pca <- data.frame(pca$x)
stest.pca <- data.frame(pca_test$x)

strain3$PC1 <- strain.pca$PC1
strain3$PC2 <- strain.pca$PC2


stest3$PC1 <- stest.pca$PC1
stest3$PC2 <- stest.pca$PC2



model3 <- glm(class ~.,family=binomial(link='logit'),data=strain3)

fitted.results3 <- predict(model3,newdata=stest3[-14],type='response')
fitted.results3 <- ifelse(fitted.results3 > 0.5,'>50K.','<=50K.')
xtab <- table(fitted.results3,stest3$class)

avgAccuracy = sum(diag(xtab)) / sum(xtab)
avgAccuracy

```
No change
